### 2020-06-17 latex/symbols.tex
Added the R-mod-T symbol and a shorthand for it.
Updated the intro to mech systems
This changes the introduction to mechanical systems, and completes the
description of Lagrangian mechanics (which may be too detailed for the
purposes of this thesis). It also adds new symbols and extra commands
for easily creating partial derivatives.

% Define a table for the symbols using the booktabs convention
\noindent
\begin{table}[h!]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{ll}
   \toprule
   Symbol & Definition \\
   \midrule
   \(\R^n\) & Real numbers in \(n\) dimensions \\
   \(q_u\) & Unactuated coordinates \\
   \(q_a\) & Actuated coordinates \\
   \(p_u\) & Conjugate of momentum to \(q_u\) \\
   \(p_a\) & Conjugate of momentum to \(q_a\) \\
   A & Some really long description to see how this works when we write long text inside the table. \\
   \bottomrule
\end{tabular}

>>>>

% Define a table for the symbols using the booktabs convention
\noindent
\begin{table}[h!]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{ll}
   \toprule
   Symbol & Definition \\
   \midrule
   \(\R^n\) & Real numbers in \(n\) dimensions. \\
   \(\R^{n \times m}\) & Real-valued matrix with \(n\) rows and \(m\) columns. \\
   \(\Rt{T}\) & Real numbers modulo \(T > 0\), with \(\Rt{\infty} = \R\). \\
   \(\mathbb{S}^1\) & The unit circle, equivalent to \(\Rt{2\pi}\). \\
   \([M]_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
   \(\dot{x}\) & Derivative of \(x(t,y)\) with respect to time \(t\). \\
   \(x'\) & Derivative of \(x(t,y)\) with respect to a non-time input \(y\). \\
   \(\mathcal{Q}\) & The configuration manifold of a mechanical system. \\
   \(q_u\) & Unactuated coordinates. \\
   \(q_a\) & Actuated coordinates. \\
   \(p_u\) & Conjugate of momentum to \(q_u\). \\
   \(p_a\) & Conjugate of momentum to \(q_a\). \\
   \bottomrule
\end{tabular}

### 2020-06-17 latex/vnhcs.tex
Updated the intro to mech systems
This changes the introduction to mechanical systems, and completes the
description of Lagrangian mechanics (which may be too detailed for the
purposes of this thesis). It also adds new symbols and extra commands
for easily creating partial derivatives.

\chapter{Development of Virtual Nonholonomic Constraints}
\section{Mechanical Systems}
\subsection{Lagrangian Formulation of Mechanical Systems}
% TODO: Describe the basics of the lagrangian formulation for mechanical systems
\subsection{Hamiltonian Formulation of Mechanical Systems}
% TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
% you can write the hamiltonian directly in certain conditions of the system
\subsection{Simply Actuated Hamiltonian Systems}
% TODO: Describe the change of coordinates to get into q_u/q_a mode and show
% that the actuator directly affects pa but not pu. Use M for simply actuated
% inertia, D for normal coordinates

>>>>

\chapter{Development of Virtual Nonholonomic Constraints}
\section{Preliminaries on Analytical Mechanics}
In classical mechanics, a mechanical system is comprised of \(N\) point masses
representing the center of mass of a physical body, where the position of each
mass is described using three cartesian coordinates (one for each spatial axis). 
The system as a whole is represented by a vector in \(\R^{3N}\), and the
dynamics of the system are computed by deriving the \(3N\) \textit{equations of
motion} (EOM) produced by Newton's second law \(F = m a\).  
While this method works for simple cases, it unfortunately proves impossible
to apply to a majority of mechanical systems since most of the forces
are not known explicitly.

Most mechanical systems are not solely a collection of point masses which can 
freely move about three-dimensional space; the masses are usually restricted
physically by some equations of constraint (EOC). 
Rather than representing the system by one vector in \(\R^{3N}\) and \(k\) EOC, 
it is often feasible to represent the position of the system using \(n\)
independent scalar-valued variables \(q_1,\ldots,q_n\) called 
\textit{generalized coordinates}, where \(n = 3N - k\) is the number of
\textit{degrees of freedom} (DOF) of the system.

Each generalized coordinate \(q_i\) takes values in \(\Rt{T_i}\), where
\(T_i = \infty\) if \(q_i\) represents a length or \(T_i = 2\pi\) if \(q_i\)
represents an angle.
It is convention to write these as a \textit{configuration} vector
\(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) where the \textit{configuration space}
\(\mathcal{Q}\) has the following structure:
\[
    \mathcal{Q} = \Rt{T_1} \times \cdot \times \Rt{T_n}
\] 
The derivative \(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n)\) of a configuration
is called a \textit{generalized velocity}, and the joint vector
\((q,\dot{q})\) is called a \textit{state} of the system.

The field of analytical mechanics provides a computational method for finding
the EOM of a system in generalized coordinates. The two most common methods for
modelling robotic systems are \textit{Lagrangian} and \textit{Hamiltonian}
mechanics.

% ---------- Lagrangian Mechanics ---------- % 
\subsection{Lagrangian Mechanics}

Lagrangian mechanics uses the kinetic energy \(T(q,\dot{q})\) and potential
energy \(P(q)\) of the system to define the Lagrangian
(\ref{eqn:lagrangian_general})
\begin{equation}\label{eqn:lagrangian_general}
    \mathcal{L}(q,\dot{q}) = T(q,\dot{q}) - P(q)
\end{equation}
In actuated mechanical systems, the EOM are described by \(n\) second-order
differential equations obtained from the \textit{Euler-Lagrange
equations} with \textit{input torques} \(\tau \in \R^m\) 
(\ref{eqn:el_eqns_general}). 
\begin{equation}\label{eqn:el_eqns_general}
    \diff{}{t}\left\{ \pdiff{\mathcal{L}}{\dot{q}_i} \right\}
    - \pdiff{\mathcal{L}}{q_i} = B_i(q) \tau
\end{equation}
Here, \(B_i: \mathcal{Q} \rightarrow \R^{1\times m}\) is an input vector 
and the matrix  \(B: \mathcal{Q} \rightarrow \R^{n \times m}\) given by
\[
    B(q) =\begin{bmatrix}
        - & B_1(q) & - \\
          & \vdots & \\
        - & B_n(q) & -
    \end{bmatrix}
\]
is the \textit{input matrix} for the system. If \(m < n\), we say the system is
\textit{underactuated} with degree of underactuation \(n - m\).

Many actuated mechanical systems have quadratic kinetic energies, so that the
Lagrangian can be written explicitly as
\[
    \mathcal{L}(q,\dot{q}) = \dot{q}\tpose D(q) \dot{q} - P(q)
\]
where the \textit{inertia matrix} \(D: \mathcal{Q} \rightarrow \R^{n\times n}\) 
is symmetric, positive definite matrix for all \(q \in \mathcal{Q}\) and the
potential function \(P : \mathcal{Q} \rightarrow \R\) is smooth. 
If this is the case, the Euler-Lagrange equations
(\ref{eqn:el_eqns_general}) reduce to (\ref{eqn:el_eqns}),
\begin{equation}\label{eqn:el_eqns}
    D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
\end{equation}
where the \textit{Coriolis matrix} \(C(q,\dot{q})\) is of the form
\[
    [C]_{i,j} = \frac{1}{2}\sum\limits_{k = 1}^n 
    \left(\pdiff{[D]_{i,j}}{q_k}  +
     \pdiff{[D]_{i,k}}{q_j} -
     \pdiff{[D]_{k,j}}{q_i}\right)\dot{q}_k
\]

% ---------- Hamiltonian Mechanics ---------- %
\subsection{Hamiltonian Mechanics}
% TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
% you can write the hamiltonian directly in certain conditions of the system

\section{Canonical Control Coordinates for Hamiltonian Systems}
% TODO: Describe the change of coordinates to get into q_u/q_a mode and show
% that the actuator directly affects pa but not pu. Use M for simply actuated
% inertia, D for normal coordinates
